---
title: "Lab 9 Base R"
author: "Hollie Pennington"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#objects and classes 

## Vectors 

### Creating a vector 

```{r}
#numeric vector 
my_vector <- vector(mode = "numeric", length = 100)
#logical vector
my_second_vector <- vector(mode = "logical", length = 20)
```

### Vector properties 


#### Length 

```{r}
#length of my vector
length(my_vector)
length(my_second_vector)
```

#### class 

#vector properties 

```{r}
# what's in my vector
class(my_vector)
class(my_second_vector)
```

#### Indexing (accessing) vectors

```{r}
# using c to create a vector 
my_vector <- c(1:50)

# what's the first element? 
my_vector[1]

# what's the last element? 
my_vector[-1]

# what's the seventh element? 
my_vector[7]

# select a range from a vector 
my_vector[3:7]
```

## Matrices 

### creating matrices 

```{r}
#Create a 10 X 10 matrix full of NA's 
my_matrix <- matrix(data = 1:100, nrow = 20, ncol = 5)
```

### Matrix properties 

#### dimensions 

```{r}
#what are my dimensions 
dim(my_matrix)
```

#### Length 

```{r}
# how many elements are there? 
length(my_matrix)
```

#### Class

```{r}
class(my_matrix)
```
#### Indexing 

```{r}
my_matrix[2,]
# index a column 

my_matrix[,3]

#index a given cell 
my_matrix[8,3]

# indexing and replace 
my_matrix[8,3] <- 100000

# replace an entire column 
my_matrix[,5] <- 10
```

# control flow 

## 'for' loops 

variable (counter), sequene, expression 

```{r}
for (i in 1:10){
  print("i")
}
```

#generating random numbers

```{r}
rnorm(n = 5, mean = 10, sd = 2)
```

#lets combine all of the above 

We will use 'numeric' to create an empty vector 
we will use 'for' to iterate along items of that empty vector 
we will use '[]' to access each a lot and fill-in that empty vector 
we will use 'rnorm' to sample values from a random normal distribution 
we will plot the trend

$$N_{t}=N_{t-1} + r\left(1 - \frac{N_{t-1}}{K}\right)$$

$$
r \sim norm(\mu = 1.5, \sigma = 0.8)
$$
## writing the code down 

```{r}
# set up model parameters 
K <- 100 
nsteps <- 100
mu <- 1.5
sigma <- 0.8 

# set up model objects 
N <- vector(mode = "numeric", length = nsteps)
N[1] <- 25

# model it 
for(t in 2:nsteps){
  r <- rnorm(n = 1, mean = mu, sd = sigma)
  N[t] <- N[t-1] + r * (1 - (N[t-1]/K))
}

#visualize it 
plot(N)
```

# Something More 

We will use 'matrix' to create an empty matrix (rows are time, columns are runs)
we will repeat the process 

```{r}
# set up model parameters 
K <- 100 
nsteps <- 100
mu <- 0.4
sigma <- 1 

# set up model objects 
N <- matrix(data = NA, nrow = nsteps, ncol = 100)
N[1,] <- 25

# model it 
for(t in 2:nsteps){
  r <- rnorm(n = 100, mean = mu, sd = sigma)
  N[t,] <- N[t-1,] + r * (1 - (N[t-1,]/K))
}

#visualize it 
matplot(N, type = "l")
```

# Bring in your freinds! 

Covert it to thinks we know about: 'ggplot2' 

```{r}
# load tidyverse
library(tidyverse)

# create date.frame 
my_data <- as_tibble(N) %>%
  mutate(time = 1:nsteps) %>% 
  gather(run, N, -time)

#visualize it 
ggplot(data = my_data, mapping = aes(x = time, y = N)) + 
  geom_line(aes(group = run), size = 0.1, alpha = 0.5) + 
  stat_summary(geom = "line", fun = mean, color = "red") + 
  theme_bw() + 
  labs(x = "Time (years)", y = "Population size (N)")

```



